A Graph Theory Based Approach to MDAO Problem Formulation

### 2013-03-21 graph_syntax.tex
wrote design vars and local vs global sections

\section{Requirements for New Graph Syntax}
The ultimate goal of the new graph--based syntax presented here is to be able to 
fully describe the fundamental problem formulation for a complex design task, 
while still being able to accomodate the more specific case when a solution 
strategy is applied to the problem formulation. In order to achieve that goal 
the graph needs to accomodate a number of constructs from MDAO problems:

@@

The relationship between these three graphs is depicted in Figs.~\ref{f:tree} and \ref{f:hourglass}. 
The tree diagram demonstrates the fact that it is generally possible to obtain 
multiple FPGs from a single maximal connectivity graph. This  may correspond to 
different down--selections of analysis codes, different connections between them, 
or both. Each down--selection shrinks the number of possible FPGs that could be reached 
untill only one is remaining. Then, from a single FPG, different PSGs may be obtained by implementing 
different solution strategies. The hourglass shape in Fig. \ref{f:hourglass} illustrates how 
the MCG gets reduced to a single FPG, then multiple possible PSGs exist to solve the problem.
In other words the FPG is obtained from the MCG by removing nodes and edges, 
and the PSG is obtained from the FPG by adding nodes and edges.

@@

\section{Syntax Definition}
In the rest of this section we present the necessary graph theory fundamentals to 
construct any of the different graphs discussed above. 
The notation used in this work is adapted from Diestel \cite{Diestel2010}. 
A \emph{graph} is a pair $G = (V,E)$ of sets such that $E \subseteq V \times V$, 
which means that the elements of $E$ are 2--element subsets of $V$. The set $V$ 
contains the \emph{vertices} or \emph{nodes} and the set $E$ contains the \emph{edges}.
For a \emph{directed graph*} we construct $E$ as a set of ordered pairs instead 
of a set of sets. Each ordered pair represents an edge starting at the node 
indicated by the first entry and directed to the node indicated by the second 
entry. Edge $e$ = $(x,y)$ may be referred to simply as $xy$ and we say $y = E(x)$.

@@

Let $I$ be a nonempty set such that for each $i \in I$ there is a corresponding set $A_i$. 
The family $\{A_i \st i \in I\}$ is an indexed family of sets with index $i$ and 
indexing set $I$, and we can say $\mathcal A = \{A_i \st i \in I\}$\cite{smith2006}. 
Then the union over this family of sets is

\begin{equation}
\bigcup_{i \in I} A_i = \bigcup_{A \in \mathcal A} A = \{x \st x \in A \txt{ for } A \in \mathcal A\}.
\end{equation}
Lastly, the cardinality of a set $B$ is the number of elements in $B$ and is denoted as $|B|$.

There are four node types:  
\begin{description}
\item[variable: ] represents scalar or array data
\item[model:] responsible for mapping inputs to outputs
\item[expression:] simple algebreic combinations of variables (such as for 
  constraint of objective functions)
\item[driver:] constrol structures capable of managing iteration
\end{description}

@@

@@

For node $v \in V$ the edges directed out are given by $E(v)$ and the edges 
directed into $v$ are given by $E^{-1}(v)$; $E(E(v))$ is denoted as $E^2(v)$, 
and likewise for additional levels. The \emph{indegree} of a node is the number 
of explicit edges directed in and is denoted as $\txt{deg}^-(v)$, and the \emph{outdegree} 
is the number of explicit edges directed out and it is denoted as $\txt{deg}^+(v)$.
We also define the \emph{upper indegree limit}

\begin{equation}
\txt{deg}_u^-(v):V \to \mathbb{N}
\end{equation} 
and the \emph{lower indegree limit}
\begin{equation}
\txt{deg}_l^-(v):V \to \mathbb{N}.
\end{equation}

These user-specified limits govern the number of explicit edges that may be directed into a 
node for a valid graph. Consider a variable node $v$ with
$\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, that node must have one, and only one
incomming explicit edge or the graph is not valid. Two possiblities exists for voilating 
the limit conditions: 
\begin{description}
  \item[hole: ] $\txt{deg}_u^-(v) < \txt{deg}_l^-(v)$. The number of incomming edges 
  is less than $\txt{deg}_l^-(v)$. 
  \item[collision: ]$\txt{deg}_u^-(v) > \txt{deg}_u^-(v)$. The number of incomming edges 
  is greater than $ \txt{deg}_u^-(v)$.
\end{description}

\section{Graph Representation of MDAO Features}

@@

Analysis tools take in a number input variables, then perform some work to calculate 
the values for their respective outputs. In an MDAO graph, this process is 
represented by an group of nodes and edges called an \emph{analysis block}, 
shown in Fig. \ref{f:analysis block}. Within an analysis block each variable 
node represents a single input or output and is connected 
to a single model node via a fixed edge. Note that in Fig. \ref{f:analysis block}, 
the analysis block contains two model nodes, with a single edge connecting them. 
This part of the graph represents the necessary calculations to map given inputs 
to proper output values. Although analysis blocks are comprised of multiple nodes and edges, since all 
the edges within them are fixed, they become fixed structures within a graph.

@@

MDAO problems require that information be passed between sets of analyses. When information from the output of one analysis 
block is passed, or connected, to the input of another analysis block, a new explicit edge is added 
connecting the two variable nodes involved in the exchange. Explicit edges are free because, unlike the edges 
within an analysis block, they could be added or removed depending on the demands of the specific problem. 
In other words, explicit edges are created by engineers linking the output of one tool to the input 
of another one.

@@

More formally, all inputs that have no incomming fixed or explicit edges are defined
as design variables. When working with a MCG or FPG, only fixed and explicit edges are allowed, 
in which case a design variable is simply a variable node with no incomming edges. For a valid PSG 
any design variable node would have at least one incomming implicit edge from a driver node.

@@

\begin{comment}
The MDO problem formulation concepts represented by nodes are given in 
Table \ref{t:node representation}, and the concepts represented by edges are 
given in Table \ref{t:edge representation}.
% \begin{itemize}
% \item An objective or constraint is indicated by a variable node with no outgoing edges and with the incoming edges being directed from only other variable nodes.
% \item A global input is a variable node with no edges directed in and with at least two edges directed out to different variable nodes.
% \item A design variable is represented by a variable node with no edges directed in and on fixed edge directed out.
% \end{itemize}
\begin{table}[h!]
 \begin{center}
  \caption{Problem formulation concept represented by a node}
  \label{t:node representation}
  \begin{tabular}{ccc} \hline 
edges directed in & edges directed out & node representation \\ \hline
only explicit edges & none & objective or constraint\\
none & at least two explicit edges & global input \\
none & one fixed edge & design variable \\ \hline
  \end{tabular}
 \end{center}
 \vspace{-15pt}
\end{table}
\begin{table}[h!]
 \begin{center}
  \caption{Problem formulation concept represented by an edge}
  \label{t:edge representation}
  \begin{tabular}{ccc} \hline 
from node & to node & edge representation \\ \hline
variable & variable & connection/passing of a variable\\
variable & model & local input \\
model & variable & local output \\
model & model & function call \\ \hline
  \end{tabular}
 \end{center}
 \vspace{-15pt}
\end{table}
%`Connection' edges connect the local outputs from one analysis block to input nodes.
\end{comment}
\subsection{Local vs Global Properties}

\begin{itemize}
  \item what can be local or global? Variables, objectives, constraints, 
  \item What does it really mean to be local vs global? 
  \item What does the graph structure of local vs global look like? 
  \item local is a frame of reference... a variable could be global when looking at two analyses, but local to the two when looking at the three
\end{itemize}

@@

\subsection{Fundamental Problem Formulation Graph}
We now define the fundamental problem formulation graph, $F=(V_F,E_F)$, as a directed graph meeting the following conditions
\begin{enumerate}
\item[(1)] $\displaystyle{I_F \subset I}$ such that the following conditions hold
\item[(2)] $\displaystyle{ \forall i \in I, \ \exists v \in V_{A_i} \txt{ with } t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v)>0 }$
%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_{A_i} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_\txt{out} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
\item[(3)] $\displaystyle{V_{\txt{in},F} = \{v \in V_\txt{in} \st \txt{deg}^+(v) > 0\} }$ ((maybe this should be 1 instead of 0 to match the previous definition of global input))
\item[(4)] $\displaystyle{V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right)}$
\item[(5)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right), \ C_F \subset C_M}$
\item[(6)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
\end{enumerate}
The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). Requirement (2) stipulates that the analysis blocks in $F$ must each have at least one local output that is beging used. Requirement (3) stipulates that only the global inputs that are being used should be included. Lastly, requirement (6) stipulates that the number of edges directed into the variable nodes must be within the lower and upper indegree limits; if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is called a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is called a \emph{collision}.

@@

\begin{enumerate}
\item The first step is to detect holes and disconnect the explicit edges following them. These explicit edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs. The set of variable nodes which are holes is created as
\begin{equation}
H = \{v \in V \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
\end{equation}
which is the set of variable nodes with few incoming edges than are allowed by the lower indegree limit.
Then the updated set of edges is created by removing the first set of explicit edges following the hole:
\begin{equation}
C_{F,1} = C_{F,0} \setminus \{(x,y) \in E \st x=E^3(v) \txt{ for } v \in H\},
\end{equation}
%deg-v is now recaculated for the new C. 
and this step is demonstrated by Fig.~\ref{f:hole}. Because removing these edges can create new holes, this step must be repeated until no more holes are found.
\begin{figure}[htb!]
	\begin{center}
	\includegraphics[width=3in]{images/analysis_block_hole}
	\end{center}
	\vspace{-20pt}
\caption{Example variable node indicating a hole.}
\label{f:hole}
\end{figure}

\item The second step is to detect collisions and to then disconnect enough explicit edges so that the collisions are resolved. The set of variable nodes which contain collisions is created as
\begin{equation}
S_\txt{nodes} = \{v \in V \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
\end{equation}
For each collision node we can construct a set contained the edges directed in. The set containing all of these sets is constructed as
\begin{equation}
S_\txt{edges} = \big \{ \{(x,y) \in E\} \st y \in S_\txt{nodes} \big \}
\end{equation}
Let $J=\{1,2,\ldots,|S_\txt{edges}|\}$ be an indexing set for $S_\txt{edges}$ such that each $S_{\txt{edges},j}$ corresponds to a set in $S_\txt{edges}$ for $j \in J$. An examle collision is shown in Fig.~\ref{f:collision} to indicate the definition of $S_{\txt{edges},j}$. We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$. Then we may construct sets of edges as
\begin{equation}
B_j = \big \{e_k \in S_{\txt{edges},j} \st k \in \{1,2,\ldots,K\} \txt{ with } K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
\end{equation}
which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper indegree limit of $v_j$. The construction of each $B_j$ corresponds to making a decision on which edges to include and which edges not to include. Then let
\begin{equation}
C_{F,2} = \{ e \in C_{F,1} \st e \in B_j \txt{ for some } j \in J\}.
\end{equation}
\begin{figure}[htb!]
	\begin{center}
	\includegraphics[width=3in]{images/analysis_block_collision}
	\end{center}
	\vspace{-20pt}
\caption{Example variable node indicating a collision.}
\label{f:collision}
\end{figure}

>>>>

\section{Requirements for New Graph Syntax}
The ultimate goal of the new graph--based syntax presented here is to be able to 
fully describe the general structure of an MDAO problem indpendently of any solution information, 
while still being able to accomodate the more specific case when a solution 
strategy is applied. In order to achieve that goal 
the graph needs to accomodate a number of constructs of MDAO problems:

@@

The relationship between these three graphs is depicted in Figs.~\ref{f:tree} and \ref{f:hourglass}. 
The tree diagram demonstrates the fact that it is generally possible to obtain 
multiple FPGs from a single maximal connectivity graph. This  may correspond to 
different down--selections of analysis codes, different connections between them, 
or both. Each down--selection shrinks the number of possible FPGs that could be reached 
untill only one is remaining. Then, from a single FPG, different PSGs may be obtained by implementing 
different solution strategies. If you consider the size of a graph to be the sum of all its
edges and nodes then the hourglass shape in Fig. \ref{f:hourglass} illustrates how
the MCG gets reduced to a single FPG, then multiple possible PSGs exist to solve the problem.
In other words the FPG is obtained from the MCG by removing nodes and edges, 
and the PSG is obtained from the FPG by adding nodes and edges.

@@

\section{Syntax Definition}
In the rest of this section we present the necessary graph theory fundamentals to 
construct any of the different graphs discussed above. 
The notation used in this work is adapted from Diestel \cite{Diestel2010}. 
A \emph{graph} is a pair $G = (V,E)$ of sets such that $E \subseteq V \times V$, 
which means that the elements of $E$ are 2--element subsets of $V$. The set $V$ 
contains the \emph{vertices} or \emph{nodes} and the set $E$ contains the \emph{edges}.
For a \emph{directed graph*} we construct $E$ as a set of ordered pairs instead 
of a set of sets. Each ordered pair represents an edge starting at the node 
indicated by the first entry and directed to the node indicated by the second 
entry. Edge $e$ = $(x,y)$ may be referred to simply as $xy$. For node $v \in V$ 
the edges directed out are given by $E(v)$ and the edges directed into $v$ are given 
by $E^{-1}(v)$; $E(E(v))$ is denoted as $E^2(v)$, and likewise for additional levels.

@@

%does this belong in the analysis block section???? 
Let $I$ be a nonempty set such that for each $i \in I$ there is a corresponding set $A_i$. 
The family $\{A_i \st i \in I\}$ is an indexed family of sets with index $i$ and 
indexing set $I$, and we can say $\mathcal A = \{A_i \st i \in I\}$\cite{smith2006}. 
Then the union over this family of sets is

\begin{equation}
\bigcup_{i \in I} A_i = \bigcup_{A \in \mathcal A} A = \{x \st x \in A \txt{ for } A \in \mathcal A\}.
\end{equation}

Lastly, the cardinality of a set $B$ is the number of elements in $B$ and is denoted as $|B|$.

\subsection{Node and Edge Types}

There are four node types:  
\begin{description}
\item[variable: ] represents scalar or array data
\item[model:] responsible for mapping inputs to outputs
\item[expression:] simple algebreic combinations of variables (such as for 
  constraint of objective functions)
\item[driver:] control logic capable of managing iteration
\end{description}

@@

\subsection{Rules for Nodes and Edges}

@@

\subsection{In-degree and Out-degree}

The \emph{in-degree} of a node is the number of fixed or explicit edges directed in and 
is denoted as $\txt{deg}^-(v)$, and the \emph{out-degree} 
is the number of fixed or explicit edges directed out and it is denoted as $\txt{deg}^+(v)$.
We also define the \emph{upper in-degree limit}

\begin{equation}
\txt{deg}_u^-(v):V \to \mathbb{N}
\end{equation} 
and the \emph{lower in-degree limit}
\begin{equation}
\txt{deg}_l^-(v):V \to \mathbb{N}.
\end{equation}

These user-specified limits govern the number of explicit edges that may be directed into a 
node for a valid graph. Consider a variable node $v$ with
$\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, that node must have one, and only one
incomming explicit edge or the graph is not valid. Two possiblities exists for voilating 
the limit conditions: 
\begin{description}
  \item[hole: ] The number of incomming edges is less than $\txt{deg}_l^-(v)$. 
    \begin{equation} \txt{deg}_u^-(v) < \txt{deg}_l^-(v) \label{e:hole} \end{equation}

  \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
    \begin{equation} \txt{deg}_u^-(v) > \txt{deg}_u^-(v) \end{equation}
\end{description}

\section{Graph Representation of MDAO Constructs}

@@

Analysis tools take in a number input variables, then perform some work to calculate 
the values for their respective outputs. In an MDAO graph, this process is 
represented by an group of nodes and edges called an \emph{analysis block}, 
shown in Fig. \ref{f:analysis block}. Within an analysis block each variable 
node represents a single input or output and is connected 
to a single model node via a fixed edge. Note that in Fig. \ref{f:analysis block}, 
the analysis block contains two model nodes, with a single edge connecting them. 
This edge represents the necessary calculations to map given inputs 
to proper output values. If constructing a weighted graph, the weight calculation edge 
provides the opportunity to computational cost. However, it is also allowed to skip 
this edge and have both incomming and outgoing edges to variables through a single model 
node for a given analysis block. Although analysis blocks are comprised of multiple nodes and edges, since all 
the edges within them are fixed, they become fixed sub-graphs within a graph.

@@

MDAO problems require that information be passed between sets of analyses. When information from the output of one analysis 
block is passed, or connected, to the input of another analysis block, a new explicit edge is added 
connecting the two variable nodes involved in the exchange. Explicit edges are free because, unlike the edges 
within an analysis block, they could be added or removed depending on the demands of the specific problem. 
In other words, explicit edges are created by engineers linking the output of one tool to the input 
of another one. Note that it is allowed for a single output to have outgoing explicit edges to multiple 
other inputs. In Fig [insert ref], an output from analysis block A is connected 
to the inputs for both blocks B and C.

[[insert graph fig here]]

@@

More formally, any input variable that can be defined as a hole by Eqn. \ref{e:hole}
would be considered a design variable in the graph. When working with a MCG or FPG, only fixed and explicit edges are allowed, 
in which case a design variable is simply a variable node with no incomming edges. For a valid PSG 
any design variable node would have at least one (but possible multiple) incomming 
implicit edge from a driver node.

@@

\begin{comment}
The MDO problem formulation concepts represented by nodes are given in 
Table \ref{t:node representation}, and the concepts represented by edges are 
given in Table \ref{t:edge representation}.
% \begin{itemize}
% \item An objective or constraint is indicated by a variable node with no outgoing edges and with the incoming edges being directed from only other variable nodes.
% \item A global input is a variable node with no edges directed in and with at least two edges directed out to different variable nodes.
% \item A design variable is represented by a variable node with no edges directed in and on fixed edge directed out.
% \end{itemize}
\begin{table}[h!]
 \begin{center}
  \caption{Problem formulation concept represented by a node}
  \label{t:node representation}
  \begin{tabular}{ccc} \hline 
edges directed in & edges directed out & node representation \\ \hline
only explicit edges & none & objective or constraint\\
none & at least two explicit edges & global input \\
none & one fixed edge & design variable \\ \hline
  \end{tabular}
 \end{center}
 \vspace{-15pt}
\end{table}
\begin{table}[h!]
 \begin{center}
  \caption{Problem formulation concept represented by an edge}
  \label{t:edge representation}
  \begin{tabular}{ccc} \hline 
from node & to node & edge representation \\ \hline
variable & variable & connection/passing of a variable\\
variable & model & local input \\
model & variable & local output \\
model & model & function call \\ \hline
  \end{tabular}
 \end{center}
 \vspace{-15pt}
\end{table}
%`Connection' edges connect the local outputs from one analysis block to input nodes.
\end{comment}
\subsection{Local vs Global Nodes}

A simple definition of a global node is any node that involves information 
from more than one analysis block. Likewise, a local node is any node 
that involves information from only a single analysis block. As stated before, variable 
nodes with fixed edges directed to model nodes (as part of an analysis block) are inherently local 
and similarly model nodes are also local. Expression and driver nodes
are not inherently local or global though. For these, their locality is determined 
by the incomming and outgoing edges they posses. Consider the graph in Fig [insert ref], 
where the expression node $g$ represents some kind of constraint for the optimziation problem. 
Since $g$ only had edges comming from analysis block $B$, it would be considered a local constraint 
on block $B$.

[[insert graph here]]

Although we've already established that variable nodes in an analysis block are local to that 
block, global design variables are a part of many MDAO problems. We've previously defined 
a design variable as any

While this simple definition is valid, it glosses over a more subtle aspect of 
some MDAO problems. Consider the graph in Fig. [insert ref], with three analysis 
blocks A, B, and C. The expression node, $g$, is a constraint that requires information 
from $A$ and $B$, so it would be considered global by the previous definition. Since $g$
has incomming edges from block $C$, it would not be global with respect to $C$. 
You could say that $g$ is local to the group $A,B$. In this sense, locality 
becomes a relative term. A given node can be local to a specific sub-graph of a large
problem. This subtle structure can become significant when determining how to partition 
an MDAO problem for use with certain MDAO architectures. Martins and Marriage 
demonstrate how Collaborative Optimization outperforms MultiDisciplinary Feasible
when they separated a problem into two sub-problems with variables, objectives, and 
constraints local to each one\cite{Marriage2008a}.

@@

\subsection{Fundamental Problem Formulation Graph}
We now define the fundamental problem formulation graph, $F=(V_F,E_F)$, as a directed graph meeting the following conditions
\begin{enumerate}
\item[(1)] $\displaystyle{I_F \subset I}$ such that the following conditions hold
\item[(2)] $\displaystyle{ \forall i \in I, \ \exists v \in V_{A_i} \txt{ with } t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v)>0 }$
%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_{A_i} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_\txt{out} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
\item[(3)] $\displaystyle{V_{\txt{in},F} = \{v \in V_\txt{in} \st \txt{deg}^+(v) > 0\} }$ ((maybe this should be 1 instead of 0 to match the previous definition of global input))
\item[(4)] $\displaystyle{V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right)}$
\item[(5)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right), \ C_F \subset C_M}$
\item[(6)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
\end{enumerate}
The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). Requirement (2) stipulates that the analysis blocks in $F$ must each have at least one local output that is beging used. Requirement (3) stipulates that only the global inputs that are being used should be included. Lastly, requirement (6) stipulates that the number of edges directed into the variable nodes must be within the lower and upper in-degree limits; if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is called a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is called a \emph{collision}.

@@

\begin{enumerate}
\item The first step is to detect holes and disconnect the explicit edges following them. These explicit edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs. The set of variable nodes which are holes is created as
\begin{equation}
H = \{v \in V \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
\end{equation}
which is the set of variable nodes with few incoming edges than are allowed by the lower in-degree limit.
Then the updated set of edges is created by removing the first set of explicit edges following the hole:
\begin{equation}
C_{F,1} = C_{F,0} \setminus \{(x,y) \in E \st x=E^3(v) \txt{ for } v \in H\},
\end{equation}
%deg-v is now recaculated for the new C. 
and this step is demonstrated by Fig.~\ref{f:hole}. Because removing these edges can create new holes, this step must be repeated until no more holes are found.
\begin{figure}[htb!]
	\begin{center}
	\includegraphics[width=3in]{images/analysis_block_hole}
	\end{center}
	\vspace{-20pt}
\caption{Example variable node indicating a hole.}
\label{f:hole}
\end{figure}

\item The second step is to detect collisions and to then disconnect enough explicit edges so that the collisions are resolved. The set of variable nodes which contain collisions is created as
\begin{equation}
S_\txt{nodes} = \{v \in V \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
\end{equation}
For each collision node we can construct a set contained the edges directed in. The set containing all of these sets is constructed as
\begin{equation}
S_\txt{edges} = \big \{ \{(x,y) \in E\} \st y \in S_\txt{nodes} \big \}
\end{equation}
Let $J=\{1,2,\ldots,|S_\txt{edges}|\}$ be an indexing set for $S_\txt{edges}$ such that each $S_{\txt{edges},j}$ corresponds to a set in $S_\txt{edges}$ for $j \in J$. An examle collision is shown in Fig.~\ref{f:collision} to indicate the definition of $S_{\txt{edges},j}$. We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$. Then we may construct sets of edges as
\begin{equation}
B_j = \big \{e_k \in S_{\txt{edges},j} \st k \in \{1,2,\ldots,K\} \txt{ with } K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
\end{equation}
which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper in-degree limit of $v_j$. The construction of each $B_j$ corresponds to making a decision on which edges to include and which edges not to include. Then let
\begin{equation}
C_{F,2} = \{ e \in C_{F,1} \st e \in B_j \txt{ for some } j \in J\}.
\end{equation}
\begin{figure}[htb!]
	\begin{center}
	\includegraphics[width=3in]{images/analysis_block_collision}
	\end{center}
	\vspace{-20pt}
\caption{Example variable node indicating a collision.}
\label{f:collision}
\end{figure}

### 2013-03-21 intro.tex
a few more re-org changes.
wrote design vars and local vs global sections

\section{Existing Graph--Based Problem Formulation Syntax}

@@

    Of the three methods, XDSM comes the closest to fully describing a problem formulation. 
    Thought it does not contain specific variable information, this could be easily added back in
    simply by putting the variables back into the graph and drawing the apprpriat connections
    between them and the relevant analysis nodes. However, it's dependence on 
    solution strategy specifics to represent couplings can not be removed without 
    stripping it of the useful information contained in an FDT. In the following sections
    we describe a new graph syntax that retains the features of XDSM, without the dependence on 
    solution strategy (although it does allow for that information to be present). Thus 
    the new syntax allows for the compelete specification of an FPF.

>>>>

\section{Existing Graph--Based Syntax}

@@

    Of the three methods, XDSM comes the closest to fully describing a problem formulation. 
    Thought it does not contain specific variable information on the diaganol of the DSM, this could be easily added back in
    simply by putting the variables back into the graph and drawing the appropriate connections
    between them and the relevant analysis nodes. However, it's dependence on 
    solution strategy specifics to represent couplings can not be removed without 
    stripping it of the useful information contained in an FDT. We propose
    a new graph syntax that retains the features of XDSM, without the dependence on 
    solution strategy (although it does allow for that information to be present). This 
    new syntax allows for the compelete specification of an MDAO problem.
